### 2016-01-05 imaging-supplement.Rnw
error rates for all of hamby

% Template for the submission to:
%   The Annals of Probability           [aop]
%   The Annals of Applied Probability   [aap]
%   The Annals of Statistics            [aos] 
%   The Annals of Applied Statistics    [aoas]
%   Stochastic Systems                  [ssy]
%
%Author: In this template, the places where you need to add information
%        (or delete line) are indicated by {???}.  Mostly the information
%        required is obvious, but some explanations are given in lines starting
%Author:
%All other lines should be ignored.  After editing, there should be
%no instances of ??? after this line.

% use option [preprint] to remove info line at bottom
% journal options: aop,aap,aos,aoas,ssy
% natbib option: authoryear
\documentclass[aoas]{imsart}

\usepackage{amsthm,amsmath,natbib}
\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue]{hyperref}

% provide arXiv number if available:
\arxiv{arXiv:0000.0000}

% put your definitions there:
\startlocaldefs
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{subcaption}
\usepackage{lineno}
\modulolinenumbers[5]

\usepackage{dcolumn}
\newcolumntype{d}[1]{D{.}{.}{#1}}

\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\eh}[1]{{\color{blue} #1}}
\newcommand{\ac}[1]{{\color{darkgreen} #1}}

\endlocaldefs

\begin{document}

\begin{frontmatter}

% "Title of the paper"
\title{Supplement to Automatic Matching of Bullet Lands}
\runtitle{Supplement: Automatic Matching of Bullet Lands}

% indicate corresponding author with \corref{}
\begin{aug}
\author{\fnms{Eric} \snm{Hare}\corref{}\ead[label=e1]{erichare@iastate.edu}\thanksref{m1}},
\author{\fnms{Heike} \snm{Hofmann}\ead[label=e2]{hofmann@iastate.edu}\thanksref{m1}}
\and
\author{\fnms{Alicia} \snm{Carriquiry}\ead[label=e3]{alicia@iastate.edu}\thanksref{m1}}

\affiliation{Iowa State University\thanksmark{m1}}

\runauthor{E. Hare et al.}
\end{aug}

\end{frontmatter}

\linenumbers
\tableofcontents
\newpage
<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
#rm(list=ls())
#wd <- getwd()
library(extrafont)
library(knitr)
imgdir <- "Figure/"
codedir <- "code"
datadir <- "app/images/Hamby252_3DX3P1of2/"

options(replace.assign=TRUE,scipen=3, digits=2)
bstats <- read.csv("data-25-25/bullet-stats.csv")

scrubPath <- function(x) {
  splits <- strsplit(as.character(x), split="/")
  last <- sapply(splits, function(x) x[length(x)])
  gsub(".x3p","", last)
}

library(RColorBrewer)
library(ggplot2)
library(scales)
library(dplyr)
library(x3pr)
library(x3prplus)
library(grid)
library(gridExtra)
library(zoo)
library(tidyr)
library(rpart)
library(rpart.plot)
library(xtable)
library(sm)

@

\section{Cylindrical Fit}\label{supp:cylindrical}
Figure~\ref{fig:fixedX} shows the side profile of a crosscut through the surface measurements of bullet 1-5 at a fixed level. The smooth line on top is a circle, with estimated radius and center. The details of this fit are given below. We then remove the fit to investigate residuals.
%
\begin{figure}[hbtp]
  \centering
<<fixedX, dependson='data', echo=FALSE, warning=FALSE, message=FALSE, fig.height=2, fig.width=6, out.width='0.5\\textwidth'>>=

cols = c(alpha("grey60", alpha=0.6), alpha("black", 0.5))

br111 <- read.x3p(paste(datadir,"Br1 Bullet 1-5.x3p", sep = "/"))
dbr111 <- fortify_x3p(br111)

pars <- data.frame(getCircle(dbr111$y, dbr111$value))
dbr111$theta <- acos((dbr111$y-pars$x0)/pars$radius)/pi*180
dbr111 <- dbr111 %>% mutate(
  xpred = cos(theta/180*pi)*pars$radius + pars$x0,
  ypred = sin(theta/180*pi)*pars$radius + pars$y0
)

qplot(data=subset(dbr111, x <= 100*1.5625^2 & x >= 99*1.5625^2), y, value, geom="line", size=I(1)) +
  geom_line(aes(x=xpred, y=ypred, group=x), 
            colour="grey30", size=0.25) +
#  ylab(expression(paste("Surface Measurements (in ",mu,m,")", sep=""))) + 
  ylab("") +
  theme_bw() + 
  theme(legend.position="bottom") #+ coord_equal()
@
\caption{\label{fig:fixedX}Side profile of the surface measurements (in $\mu m$) of a bullet land across a fixed $x$ value. Note that the global features dominate any deviations, corresponding to the individual characteristics of striation marks.}
\end{figure}

%One reasonable approach is to fit a circle to the curve and compute the residuals from this fit.

Assume that $n$ data points are given in the form of data tuples $(x_1, y_1), (x_2, y_2), ..., (x_n, y_n)$ that are (approximately) located on a circle. We want to estimate the location of the center and radius of the best fitting circle using a least squares approach.

We minimize the following expression:
\begin{equation}\label{eq:circle}
D = \sum_{i=1}^n \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right)^2,
\end{equation}
by differentiating $D$ with respect to $r, a,$ and $b$:
let us assume that $x_i$ and $y_i$ are centered (i.e. $\sum x_i = \sum_i y_i = 0$). Note, if they are not, make a note of the current means, subtract them now and add them to $(\hat{a}, \hat{b})$ at the end.

\noindent
The  derivate of $D$ with respect to $r$ is:
\begin{eqnarray*}
\frac{d}{dr} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 r = \\
&=& 4 r \left( n r^2 - \sum_i (x_i-a)^2 - \sum_i(y_i-b)^2 \right).
\end{eqnarray*}
At the minimum:
\begin{equation}\label{eq:rmin}
\frac{d}{dr} D = 0 \stackrel{r \neq 0}{\iff} nr^2  = \sum_i (x_i-a)^2 + \sum_i(y_i-b)^2.
\end{equation}
%
%
The  derivative of $D$ with respect to $a$ is:
\begin{eqnarray*}
\frac{d}{da} D &=& 2 \sum_i \left( r^2 - (x_i-a)^2 - (y_i-b)^2 \right) 2 (x_i - a) = \\
&=& -4 \left[ a \cdot nr^2 + \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right].
\end{eqnarray*}
Using (\ref{eq:rmin}) for $nr^2$  in the equation above we get:
\begin{eqnarray*}
\frac{d}{da} D &=& -4 \left[  \sum_i a(x_i-a)^2 +  \sum_i a(y_i-b)^2  + \right. \\
&& \phantom{-4 \ \ } \left . \sum_i (x_i - a)^3  + \sum_i (x_i - a) (y_i - b)^2 \right]  = \\
&=& -4 \left[ \sum_i (x_i-a)^2 (a + x_i - a)  + \right.\\
&& \phantom{-4 \ \ } \left .\sum_i (x_i - a + a) (y_i - b)^2 \right] = \\
&=& -4 \left[ \sum_i (x_i-a)^2 x_i   + \sum_i x_i  (y_i - b)^2 \right] 
\stackrel{\begin{array}{c}\sum_i x_i = 0\\
\sum_i y_i = 0\end{array}}{=} \\
&=& -4 \left[ \sum_i x_i^3   + \sum_i x_i y_i^2  - 2a s_{xx} - 2b s_{xy} \right],
\end{eqnarray*}
where $s_{xx} = \sum_i x_i^2, s_{xy} = \sum_i x_i y_i$ and $s_{yy} = \sum_i y_i^2$.

\noindent
Likewise, we get for the derivative of $D$ with respect to $b$:
\begin{eqnarray*}
\frac{d}{db} D &=& -4 \left[ \sum_i y_i^3   + \sum_i x_i^2 y_i - 2a s_{xy} - 2b s_{yy} \right].
\end{eqnarray*}
To find the minimum we therefore get a system of two linear equations in $a$ and $b$:
\begin{eqnarray*}
2 s_{xx} a + 2 s_{xy} b = c_1 && \text{ with } c_1 = \sum_i x_i^3 + x_i y_i^2 \\
2 s_{xy} a + 2 s_{yy} b = c_2 &&\text{ with } c_2 = \sum_i x_i^2 y_i + y_i^3.
\end{eqnarray*}
The solution to the system is:
\begin{eqnarray*}
\hat{a} &=& \frac{c_1 s_{yy} - c_2 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2},\\
\hat{b} &=& \frac{c_2 s_{xx} - c_1 s_{xy}}{2 s_{xx} s_{yy} - 2 s_{xy}^2}, \text{ and}\\
\hat{r^2} &=& \frac{1}{n}s_{xx} + \frac{1}{n}s_{yy} + \hat{a}^2 + \hat{b}^2.
\end{eqnarray*}

The scatterplot in Figure~\ref{fig:residual} shows the residuals of such a fit.
In this instance, the radius is estimated as $\hat{r} = \Sexpr{pars$radius}\mu m = \Sexpr{pars$radius/1000}mm$ and the land covers about \Sexpr{diff(range(dbr111$theta))} degrees.  Both of these estimates are consistent with a 9 mm bullet fired by a Ruger P-85.
The residuals are dominated, as expected, by the grooves, which show up as large positive residuals. For a surface cross cut at $x = 100$ there is a residual circular structure that does not show up for all cross sections.

\begin{figure}[hbtp]
  \centering
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:residuala}Residual structure at $x = 1.5625$ (bottom of the bullet).}
<<residual2, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
qplot(data=subset(dbr111, x <= 1.5625), y, value-ypred, #colour=factor(x),
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  theme_bw() + 
  geom_hline(yintercept = 0, colour="grey50") +
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}    
\begin{subfigure}[b]{.49\textwidth}\centering
\caption{\label{fig:residualb} Residual structure at $x = 100.00$}
<<residual, dependson='fixedX', echo=FALSE, warning=FALSE, fig.height=3, out.width='\\textwidth'>>=
#qplot(data=subset(dbr111, x <= 80*1.5625^2 & x >=75*1.5625^2), y, value-ypred,
qplot(data=subset(dbr111, x == 100), y, value-ypred,  #colour=factor(x), 
      geom="line", size=I(1)) +
#  scale_colour_brewer("x", palette="Paired") + 
  geom_hline(yintercept = 0, colour="grey50") +
  theme_bw() + 
  ylab(expression(paste("Residuals (in ",mu,"m)", sep=""))) + 
  theme(legend.position="bottom")
@
\end{subfigure}
\caption{\label{fig:residual} Residual structure of circular fits at two different cross sections. Both residual plots show systematic structures, indicating that a circular fit is not entirely appropriate.}
\end{figure}

A single cylinder as a fit is unlikely to be a particularly good fit, because there seem to be quite massive deformations in the vertical direction. Even when we fit a circle to each cross section of the bullet, as in Figure~\ref{fig:circlefits}, this does not address all of these issues. While the wider circumference at the base of the bullet can be resolved by individual circular fits, the systematic residual structure in Figure~\ref{fig:residualb} stays the same.

<<bullet1, echo=FALSE, cache=TRUE>>=
db1 <- NULL
for (i in 1:6) {
  bname <- sprintf(file.path(datadir, "Br1 Bullet 1-%d.x3p"), i)
  dbi <- fortify_x3p(read.x3p(bname))
  dbi$part <- i
  db1 <- rbind(db1, dbi)
}

db1 <- db1 %>% group_by(part, x) %>% do (
    data.frame(., predCircle(.$y, .$value))
  )
@

<<bullet2, echo=FALSE, cache=TRUE>>=
db2 <- NULL
for (i in 1:6) {
  bname <- sprintf(file.path(datadir, "Br1 Bullet 2-%d.x3p"), i)
  dbi <- fortify_x3p(read.x3p(bname))
  dbi$part <- i
  db2 <- rbind(db2, dbi)
}

db2 <- db2 %>% group_by(part, x) %>% do (
    data.frame(., predCircle(.$y, .$value))
  )

@

\begin{figure}[hbtp]
  \centering
<<circlefits, echo=FALSE, fig.width=10, fig.height=5, out.width='\\textwidth', warning=FALSE>>=
db2$land <- db2$part
db1$land <- db1$part
qplot(y, resid, data=subset(db2, x == 100), #colour=factor(x), 
      geom="line", size=I(.75), colour=I("grey70")) + 
  facet_wrap(~land, ncol=3, labeller="label_both") + 
  scale_colour_brewer(palette="Paired") +
  theme_bw() + theme(legend.position="bottom") + 
  geom_line(aes(y, resid, group = x), colour="black", size=.75, alpha=0.5,
             data = filter(db1, land==5, x == 100)[,c("y", "resid", "x")]) +
  ylab("Residuals from circular fit") +
  ggtitle("Bullet 1-5 in black")
@
\caption{Circular fit for each cross section of bullet 2, with bullet 1-5 overlaid.\label{fig:circlefits} Bullet 1-5 matches best with bullet 2-1.}
\end{figure}

\newpage
\section{Assessing cross-correlation between signatures at multiple cross sections}\label{supp:ccf}

Figure~\ref{fig:crosscuts} shows a sequence of cross sections for bullet 1-5 (barrel 1) that are taken 50$\mu m$ apart, between 150$\mu m$  and 400$\mu m$. These are compared to a cross section at 100$\mu m$. Initially this comparison constitutes an almost perfect match between the two cross sections. However, the match quickly deteriorates with increasing distance to non-matches at a cross section of $x = 300\mu m$ or higher.  Only if the cross section is within 150$\mu m$ do we get a good visual match even when we know that the same bullet surface is being used. 
Given that we have to expect some variation in nominally the same cross section levels due to manual alignments in microscopes, we should take cross section levels into account in the automatic matching routine by evaluating matches at several cross sections. 
\begin{figure}[hbtp]
  \centering
<<crosscuts-vary, echo=FALSE, fig.width = 12, fig.height = 7, out.width = '\\textwidth', cache=TRUE, warning = FALSE>>=
paths <- file.path(datadir, dir(datadir))
crosscuts <- seq(100, 400, by = 50)
lof <- processBullets(paths[5], x = crosscuts)
lof$bullet <- paste(lof$bullet, lof$x)

reslist <- lapply(crosscuts[-1], function(cc) {
#  browser()
  b2 <- subset(lof, x %in% c(cc, 100))
  lofX <- bulletSmooth(b2)
  bAlign = bulletAlign(lofX)
  lofX <- bAlign$bullet
    b12 <- unique(b2$bullet)
  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = 25)
  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = 25)

#  threshold <- bulletPickThreshold(lofX, thresholds = seq(0.3, 1.5, by = 0.05))
#  lines <- striation_identify(lofX, threshold = threshold)
  peaks1$lines$bullet <- b12[1]
  peaks2$lines$bullet <- b12[2]
  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)

  maxCMS <- maxCMS(lines$match==TRUE)
  list(maxCMS = maxCMS, ccf = bAlign$ccf, lines=lines, bullets=lofX)
})

ccfs <- sapply(reslist, function(res) res$ccf)

lop <- lapply(reslist, function(res) {
ggplot() +
  theme_bw() + 
  geom_rect(aes(xmin=xmin, xmax=xmax, fill=factor(type)), show.legend=FALSE, ymin=-6, ymax=5, data=res$lines, alpha=0.2) +
  geom_line(aes(x = y, y = l30, linetype=bullet),   data = res$bullets, alpha=0.6) +
  scale_colour_brewer("", palette="Set1", na.value=alpha("grey50", alpha=0.5)) +
  scale_linetype_discrete("") +
  scale_fill_brewer("", palette="Set2", na.value=alpha("grey50", alpha=0.5)) +
  theme(legend.position = c(1,1.2), legend.justification=c(1,1),
        legend.background = element_rect(fill=alpha('white', 0.4))) + 
  ylim(c(-6,6)) +
  geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(res$lines, !match)) +
  geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(res$lines, match)) +
    ylab("") + xlab("")
})

grid.arrange(lop[[1]], lop[[2]], lop[[3]], lop[[4]], lop[[5]], lop[[6]],
             ncol = 2)
@
\caption{\label{fig:crosscuts}Overview of the variations in the surface along different cross sections. Cross section at $x = 100$ is compared to cross sections every 50$\mu m$. With every step away from the original cross section site, the number of differences between the cross sections increases, and the number of maximum CMS decreases from initially 14 to  3 or fewer at $x = 300\mu m$ and higher. }
\end{figure}

\newpage
\section{Signature intensities}\label{supp:bulletbottom}
<<setup-signatures, echo=FALSE>>=
library(x3pr)
library(x3prplus)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(zoo)

knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"

###############
# can we identify the barrels the unknown bullets came from?

new_processBullets <- function(bullet, x = 100) {
  crosscuts <- unique(fortify_x3p(bullet)$x)
  crosscuts <- crosscuts[crosscuts >= min(x)]
  crosscuts <- crosscuts[crosscuts <= max(x)]
  if (length(x) > 2) crosscuts <- crosscuts[crosscuts %in% x]

  list_of_fits <- lapply(crosscuts, function(x) {
    br111 <- get_crosscut(path = NULL, x = x, bullet = bullet)
    br111.groove <- get_grooves(br111)
    fit_loess(br111, br111.groove)$resid$data
  })
  lof <- list_of_fits %>% bind_rows

  cbind(lof, bullet = bullet$path)
}

new_bulletGetMaxCMSXXX <- function(lof1, lof2, span=35) {
  lof <- rbind(lof1, lof2)
  bAlign = bulletAlign(lof)
  lofX <- bAlign$bullet

  b12 <- unique(lof$bullet)
  peaks1 <- get_peaks(subset(lofX, bullet==b12[1]), smoothfactor = span)
  peaks2 <- get_peaks(subset(lofX, bullet == b12[2]), smoothfactor = span)

  #qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
  #    theme_bw() +
  #    geom_rect(data=peaks1$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25)) +
  #    geom_rect(data=peaks2$lines, aes(xmin=xmin, xmax=xmax, fill=factor(type)), ymin=-5, ymax=5, inherit.aes = FALSE, alpha=I(0.25))

  peaks1$lines$bullet <- b12[1]
  peaks2$lines$bullet <- b12[2]

  lines <- striation_identifyXXX(peaks1$lines, peaks2$lines)

  #   p <- qplot(x=y, y=resid, geom="line", colour=bullet, data=lofX, group=bullet) +
  #     theme_bw() +
  #     geom_rect(data=lines, aes(xmin=xmin, xmax=xmax, fill = factor(type)),  ymin=-6, ymax=6, inherit.aes = FALSE, alpha=I(0.25)) +
  #     ylim(c(-6,6)) +
  #     geom_text(aes(x = meany), y= -5.5, label= "x", data = subset(lines, !match), inherit.aes = FALSE) +
  #     geom_text(aes(x = meany), y= -5.5, label= "o", data = subset(lines, match), inherit.aes = FALSE)

  maxCMS <- maxCMS(lines$match==TRUE)
  list(maxCMS = maxCMS, ccf = bAlign$ccf, lag=bAlign$lag, lines=lines, bullets=lofX)
}

# match unknown land using crosscuts
ccs <- read.csv("csvs/crosscuts-25.csv")
all_bullets <- lapply(as.character(ccs$path), function(x) {
  result <- read.x3p(x)
  result[[3]] <- x
  names(result)[3] <- "path"

  return(result)
})

knowns <- all_bullets[1:120]
unknowns <- all_bullets[121:210]

if (!file.exists("csvs/crosscut-sd.csv")) {
  bullets_processed <- lapply(all_bullets, function(bul) {
    cat("Computing processed bullet", basename(bul$path), "\n")
    crosscuts <- 25*(1:20)
    dframe <- new_processBullets(bullet = bul, x = crosscuts)
    dframe$bullet <- with(dframe, paste(bullet, x))
    dframe
  })
  names(bullets_processed) <- as.character(ccs$path)

  bullets_smoothed <- bullets_processed %>% bind_rows %>% bulletSmooth

  stats <- bullets_smoothed %>% group_by(bullet, x) %>% summarize(
    sd = sd(l30, na.rm=T)
  )
  splits <- strsplit(stats$bullet, split=".", fixed=TRUE)
  bnames <- sapply(splits, function(x) paste(x[1],".x3p", sep=""))
  stats$bullet <- bnames

  write.csv(stats, "csvs/crosscut-sd.csv", row.names=FALSE)
} else {
  stats <- read.csv("csvs/crosscut-sd.csv")
}  
@
%
Figure~\ref{fig:overview} shows an overview of the signatures across different crossucts on a single bullet. 
\begin{figure}[hbtp]
<<one-bulletland-sd, echo=FALSE, dependson='setup-signatures', fig.height=5, fig.width=8, out.width='0.7\\textwidth', warning=FALSE>>=
k <- 1
x <- as.character(ccs$path)[k]
land <- read.x3p(x)
land[[3]] <- x
names(land)[3] <- "path"

crosscuts <- 25*(1:20)
dframe <- new_processBullets(bullet = land, x = crosscuts)
dframe$bullet <- with(dframe, paste(bullet, x))

bullet <- dframe %>% bind_rows %>% bulletSmooth 
subbullet <- subset(bullet, x >= ccs$cc[k])
qplot(x=y, geom="line", y=l30, colour=x, group=x, data=subbullet) + theme_bw() +
  scale_colour_gradient("Crosscuts") +
  theme(legend.position="bottom") +
  ylab("Residuals from loess fit")
@
\caption{\label{fig:overview}Signatures of the same bullet at different crosscuts. With increasing crosscut level, peaks and valleys are less pronounced, resulting in a smaller standard deviation.}
\end{figure}
As crosscut levels increase individual characteristics become less distinctive, making true matches to other bullets harder. The pattern of decreasing peaks and valleys is generally true for  bullet lands, as can be seen in Figure~\ref{fig:sds}. 
%
\begin{figure}[hbtp]
\centering
<<crosscuts-sd, echo=FALSE, fig.height=5, fig.width=8, out.width='.7\\textwidth', dependson='setup-signatures'>>=
qplot(x, sd, data=stats, geom="line", group=bullet, alpha=I(0.5)) + theme_bw() +
  geom_smooth(group=1) + 
  xlab("Crosscut location") +
  ylab("Standard deviation")
@
\caption{\label{fig:sds}Standard deviation reduces as crosscut levels increase.}
\end{figure}
Figure~\ref{fig:sds} shows that the amount of standard deviation of a signature decreases on average for all bullet lands as crosscut levels increase.
This makes standard deviation of a signature one measure to quantify the extent to which a signature is expressed. This implies that for identifying matches we should use the lowest crosscut level to extract a bullet's signature once a stable surface region is detected. This is in accordance with current standard practice \citep{afte:1992}.

\newpage
\section{Complete evaluation of the Hamby study}\label{supp:extended}
\hh{One way to expand the use of the James Hamby study is to not only match all of the unknown bullet lands against the known bullet lands, but to compare every land against every other land. This effectively doubles the number of comparisons from 10,384 pairwise comparisons of useable bullet lands to 21,115 ($= (118+88)\cdot 205/2$) comparisons by adding another 10,731 bullet land comparisons made up of known-to-known and unknown-to-unkown comparisons. }

\hh{When we predict the new 10,731 comparisons using the random forest based on the previously considered 10,384 known-unknown comparisons, we encounter 18 false negatives and 9 false positives, corresponding to an actual false error rate of 0.19 and a false positive rate of 0.00085, which is close to the estimated error rates of 0.2267 and 0.0002. }

\hh{However, if we use all of the available comparisons to fit another random forest of 300 trees, the de-facto error rates for false positives and false negatives are again at 0. The estimated OOB error rates are 0.00024 for the false positive rate and 0.22180 for the false negative rate.}
<<allhamby, echo=FALSE, message=FALSE, warning=FALSE>>=
knowndatadir <- "app/images/Hamby252_3DX3P1of2/"
knowns <- dir(path=knowndatadir)
knowns <- gsub(".x3p", "", knowns)

unknowndatadir <- "app/images/Hamby252_3DX3P2of2/"
unknowns <- dir(path=unknowndatadir)
unknowns <- gsub(".x3p", "", unknowns)

#bstats <- read.csv("data-25-25/bullet-stats.csv")
#flagged <- c("Br6 Bullet 2-1", "Br9 Bullet 2-4", "Ukn Bullet B-2", "Ukn Bullet Q-4")
includesVar <- setdiff(names(ballstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "id", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km", "forestsmall",
                                        "left_cms", "right_cms", "left_noncms", "right_noncms"))

library(randomForest)
#set.seed(20151202)
#rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=100)

ballstats <- read.csv("data-new-all-25-25/bullet-stats-single.csv")
# names(ballstats)[10] <- "num.nonmatches"
# names(ballstats)[11] <- "CNMS"
# names(ballstats)[16] <- "S"
# names(ballstats)[17] <- "CMS"
#ballstats$forestsmall <- predict(rtrees, newdata=ballstats, type="prob")[,2]
###########

ballstats$insmall <- with(ballstats, ((b1 %in% knowns) & (b2 %in% unknowns)) | ((b2 %in% knowns) & (b1 %in% unknowns)))

set.seed(20160512)
rtrees2 <- randomForest(factor(match)~., data=subset(ballstats, !flagged & insmall)[,includesVar], ntree=300)

ballstats$forestsmall <- predict(rtrees2, newdata=ballstats, type="prob")[,2]

# xtabs(~(forestsmall>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])

set.seed(20151201+i); i+1
rtrees3 <- randomForest(factor(match)~., data=subset(ballstats, !flagged)[,includesVar], ntree=300); rtrees3
ballstats$forest <- predict(rtrees3, newdata=ballstats, type="prob")[,2]
xtabs(~(forest>0.5)+match+insmall, data=ballstats[!ballstats$flagged,])
qplot(forest, data=subset(ballstats, !flagged), geom="jitter", y=match)
@
% 9/(10628+9)
% 18/(76+18)

\hh{Show aligned signatures for all ten barrels. }

Figure~\ref{fig:aligned} gives an overview of all the signatures from bullet lands in the Hamby study aligned by barrel. Three to five bullets were fired from each barrel. The figure gives us both some insight into how well signatures match and how consistent individual characteristics are impregnated on bullets fired from each of the barrels. Signatures for some lands match remarkably well - such as land 5 from barrel 1, whereas all lands from barrel 5 show some variability both in the location and depths of peaks and valleys.
\begin{figure}[hbtp]
<<aligned, echo=FALSE, warning=FALSE, fig.width=12, fig.height=15, out.width='\\textwidth', cache=TRUE>>=
matches <- read.csv("csvs/matches.csv", header = FALSE)
matches$barrel <- rep(1:10, each=6)
matches$id <- 1:nrow(matches)
library(reshape2)
mm <- melt(matches, id.var=c("barrel", "id"), na.rm=TRUE)
mm <- subset(mm, value != "")
mm$value <- gsub("Br.* Bullet ", "", mm$value)
mm$prefix <- sprintf("Br%d Bullet ", mm$barrel)
unkns <- grep("[A-Z]",mm$value)
mm$prefix[unkns] <- "Ukn Bullet "
mm$path <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P1of2//"
mm$path[unkns] <- "~/papers/2015-imaging-paper/app/images/Hamby252_3DX3P2of2//"
mm$path <-   with(mm, paste0(path, prefix, value, ".x3p"))

ccs <- read.csv("csvs/crosscuts-25-new.csv")
mm <- merge(mm, ccs, by="path", all=TRUE)
mm <- subset(mm, !is.na(mm$cc))

crosscuts <- plyr::ldply(
  1:nrow(mm), 
  function(i) {
    dframe <- get_crosscut(mm$path[i], mm$cc[i])
    groove <- get_grooves(dframe)
    data.frame(mm[i,], fit_loess(dframe, groove)$resid$data)
  })

crossSmooth <- bulletSmooth(crosscuts)

alignme <- function(data) {
#  browser()
  data$bullet <- as.character(data$bullet)
  bullets <- unique(data$bullet)
  b1 <- bullets[1]
  lofs <- NULL
  for (i in bullets[-1]) {
    lof <- subset(data, bullet %in% c(b1, i))
    lofs <- rbind(lofs, bulletAlign(lof)$bullet)
  }
  lofs <- unique(lofs)
  lofs
}

aligned <- crossSmooth %>% group_by(id) %>% do(
  data.frame(alignme(.))
)
aligned$land <- (aligned$id %% 6)
aligned$land[aligned$land==0] <- 6
qplot(x=y, y=l30, data=aligned, geom="line", group=bullet) + ylim(c(-7,7)) + facet_grid(facets=barrel~land, labeller="label_both") + theme_bw() +
  xlab("") + ylab("Signature") + 
  theme(plot.margin=unit(c(0,0,0,0), unit="cm"))
@
\caption{\label{fig:aligned}Overview of aligned signatures for all bullet lands for the ten barrels of the Hamby study. }
\end{figure}

\newpage
\section{Table of feature importance}\label{supp:randomforest}

<<setup-rtrees, echo=FALSE, message=FALSE>>=
bstats <- read.csv("data-25-25/bullet-stats.csv", stringsAsFactors = FALSE)

library(randomForest)
set.seed(20151202)
includesVar <- setdiff(names(bstats), c("b1", "b2", "data", "resID", "id.x", "id.y", "pred", "forest", "bullet", "span", "crosscutdist", "flagged", "km"))

rtrees <- randomForest(factor(match)~., data=subset(bstats, !flagged)[,includesVar], ntree=300)
@

\begin{table}[tbhp]
\caption{\label{tab:importance}Table of  features derived from bullet image ordered by importance in predicting matches. Importance is measured in terms of mean decrease in gini index when including the variable in a decision tree. Averages (and standard deviations) for known matches (KM) and known non-matches (KNM) are shown in the last four columns.}
\centering
<<importance, echo=FALSE, results='asis', warning=FALSE>>=
imp <- data.frame(importance(rtrees))
imp$Variable <- row.names(imp)
imp <- imp[order(-imp$MeanDecreaseGini),]
names(imp)[1] <- "Importance"
imp <- imp[,2:1]
imp$meanMatch <- sapply(imp$Variable, function(var) {
  mean(subset(bstats, match)[,var])
})
imp$sdMatch <- sapply(imp$Variable, function(var) {
  sd(subset(bstats, match)[,var])
})
imp$meanNonMatch <- sapply(imp$Variable, function(var) {
  mean(subset(bstats, !match)[,var])
})
imp$sdNonMatch <- sapply(imp$Variable, function(var) {
  sd(subset(bstats, !match)[,var])
})
imp$meanMatch <- round(imp$meanMatch,1)
imp$sdMatch <- sprintf("(%5.2f)", imp$sdMatch)
imp$meanNonMatch <- round(imp$meanNonMatch,1)
imp$sdNonMatch <- sprintf("(%7.2f)", imp$sdNonMatch)
row.names(imp) <- 1:nrow(imp)
imp$Variable[c(2,4)] <- c("#matches", "#non-matches")
names(imp)[-(1:2)] <- c("KM", "(sd)", "KNM", "(sd)")
print(xtable(imp, align="clrrd{3.2}rd{3.2}", digits=c(0,0,1,1,2,1,2)), 
      floating = FALSE)
@
\end{table}

\bibliographystyle{imsart-nameyear.bst}
%\biboptions{authoryear}
\bibliography{references}

\end{document}