### 2016-04-07 mean-profiles.Rmd
bitof cleaning up in the script

---
title: "extracting mean profiles"
author: "Heike Hofmann"
date: "April 6, 2016"
output: html_document
---

Extracting the mean profile is based on the really nice paper by Chu et al (2010).

```{r, echo=FALSE, warning=FALSE}

library(x3pr)
library(x3prplus)
library(ggplot2)
library(dplyr)

br115 <- "~/CSAFE/Bullets/Hamby252_3DX3P1of2/Br1 Bullet 2-1.x3p"
im115 <- fortify_x3p(read.x3pplus(br115))
qplot(y, x, fill=value, data=im115, geom="tile")
ggsave(file="im115.pdf")

gg115 <- processBullets(read.x3pplus(br115), x=unique(im115$x))
qplot(y, x, fill=resid, data=gg115, geom="tile") + scale_fill_gradient2(limits=c(-7,7))
ggsave(file="im115-resid.pdf")

# get the 'twist'
xs <- unique(gg115$x)
twist <- NULL
ccf <- NULL
aligned <- list()

for (i in seq_along(xs)[-1]) {
  profiles <- subset(gg115, x %in% xs[c(i,i-1)])  
  profiles$bullet <- sprintf("br1-1-5-%s",profiles$x)
  aligned[[i]] <- bulletAlign(profiles, value="resid")
  twist <- c(twist, aligned[[i]]$lag)
  ccf <- c(ccf, aligned[[i]]$ccf)
}

#qplot(xs, c(0, twist)) + ylim(c(-10,10))
#qplot(xs, c(1, ccf)) + geom_hline(yintercept=0.95, colour="red")
qplot(xs, cumsum(c(0, twist))) +ylim(c(700,1100))
dframe <- data.frame(x = xs, twist=cumsum(c(0, twist)), ccf = c(0,ccf))
qplot(x, twist, data=subset(dframe, between(x, 220, 600))) +geom_smooth(method="lm")

summary(m115 <- lm(twist~x, data=subset(dframe, between(x, 220, 600))))

gg115 <- merge(gg115, dframe, by="x")
qplot(y-twist, x, fill=resid, data=gg115, geom="tile") + scale_fill_gradient2(limits=c(-7,7))
ggsave(file="im115-untwist.pdf")

gg115$modeltwist <- predict(m115, newdata=gg115)
# round to precision of y:
gg115$roundtwist <- 1.5625*round(gg115$modeltwist/1.5625, 0)

qplot(y-roundtwist, x, fill=resid, data=gg115, geom="tile") + scale_fill_gradient2(limits=c(-7,7))
ggsave(file="im115-roundtwist.pdf")

qplot(x, ccf, data=dframe) + geom_hline(yintercept=0.975)

mean_profile_all <- gg115 %>% filter(ccf > .975) %>% group_by(y) %>%
  summarize(
    resid = mean(resid),
    type="mean-all"
  )

mean_high_200 <- gg115 %>% filter(ccf > .975, between(x,50,500)) %>% group_by(y) %>%
  summarize(
    resid = mean(resid),
    type="mean-high-200"
  )

mean_low_200 <- gg115 %>% filter(ccf > .975, between(x,50,375)) %>% group_by(y) %>%
  summarize(
    resid = mean(resid),
    type="mean-low-200"
  )

profiles <- gg115 %>% filter(x==100) %>% select(y, resid) %>%  transform(type="profile-at-100")
profiles <- rbind(profiles, mean_profile_all, 
                 mean_high_200, mean_low_200)
```

in black: average of all, in blue average based on profiles below 500, in green average based on profiles below 300, in red profile at 100.
```{r, echo=FALSE}
qplot(y, resid, data=profiles, geom="line", group=type, colour=type) + theme_bw() + ylim(c(-5,5))
```