### 2016-07-13 code/filtering.Rmd
The new hamby code

---
title: "Testing Filtering Parameters"
output: html_notebook
---

```{r, message=FALSE}
library(x3pr)
library(x3prplus)
library(ggplot2)
library(smoother)
library(changepoint)

mybullet <- read.x3pplus("~/GitHub/imaging-paper/app/images/Hamby252_3DX3P1of2/Br1 Bullet 1-1.x3p")
```

```{r}
mycc <- get_crosscut(x = 100, bullet = mybullet)
my.groove <- get_grooves(mycc)
my.loess <- fit_loess(mycc, my.groove)
qplot(y, resid, data = my.loess$data, geom = "line") + theme_bw()
```

```{r}
my.loess$data$smoothed_first <- smth.gaussian(my.loess$data$resid, window = 16)
my.loess$data$smoothed <- my.loess$data$smoothed_first - smth.gaussian(my.loess$data$resid, window = 160)
my.inds <- which(diff(sign(diff(my.loess$data$smoothed))) == -2) + 1
my.inds2 <- which(diff(sign(diff(my.loess$data$smoothed))) == 2) + 1

length(my.inds) + length(my.inds2)

qplot(y, smoothed, data = my.loess$data, geom = "line") + theme_bw() +
    geom_vline(xintercept = my.loess$data$y[my.inds], colour = "red") +
    geom_vline(xintercept = my.loess$data$y[my.inds2], colour = "blue")
```